---
title: "Peer assigment 1"
author: "rpowal"
date: "Sunday, March 15, 2015"
output: html_document
---

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.



####Load the data and convert to date
```{r load}
setwd('C:\\Users\\RP86831\\Desktop\\R\\reproducible\\peer1')
activity<-read.csv('activity.csv')
activity$date<- as.Date(as.character(activity$date), '%Y-%m-%d')
```


####What is mean total number of steps taken per day?
```{r steps, fig.height=4, fig.width=5}
agg<-with(activity[!is.na(activity$steps),], aggregate(steps, list(date=date),sum))
library(ggplot2)
qplot(agg$x, binwidth=3000, geom='histogram', xlab='Average steps number')
mean(agg$x)
median(agg$x)
```


####What is the average daily activity pattern?
```{r tseries, fig.height=4, fig.width=5}
interval <- with(activity[!is.na(activity$steps),], aggregate(steps, list(interval=interval),mean))
qplot(x=interval$interval, y=interval$x, geom='line', xlab='Interval', ylab='Avg. no. of steps')
with(interval, interval[x==max(x)])
```


####Imputing missing values
```{r na, fig.height=4, fig.width=5}
sum(is.na(activity$steps))
filled<-merge(x=activity, y=interval, by="interval")
filled$steps<-ifelse(is.na(filled$steps), filled$x, filled$steps)

aggfilled<-with(filled, aggregate(steps, list(date=date),sum))
qplot(aggfilled$x, binwidth=3000, geom='histogram', xlab='Average steps number')
mean(aggfilled$x)
median(aggfilled$x)

```

After filling blanks total number of steps is higher. Average number of steps and median is not significantly different. 


####Are there differences in activity patterns between weekdays and weekends?
```{r weekday, fig.height=4, fig.width=8}
workd <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
filled$wd <-  factor((weekdays(filled$date) %in% workd)+1L,
                    levels=1:2, labels=c('weekend', 'weekday'))
interval_wd <- with(filled, aggregate(steps, list(interval=interval, wd=wd),mean))
qplot(x=interval, y=x, data=interval_wd, geom='line', xlab='Interval', ylab='Avg. no. of steps', facets= .~wd)

```